Under-ice acoustic navigation using real-time model-aided range estimation

### 2021-12-22 rebuttal/rebuttal-letter/trunk-rebuttal.tex

% table of contents
\renewcommand{\contentsname}{}
\begingroup
\let\pagebreak\relax
\setcounter{tocdepth}{1}
\tableofcontents
\endgroup

\section*{Cover Letter}
Dear Editors,

We thank the reviewers and editor for their thoughtful feedback and comments in improving this work.
We have refashioned the abstract and introduction to better elucidate the motivation for this work as a model-aided component for LBL navigation in total under-ice conditions, which could be further generalized for any acoustic environment.
Based on several suggestions, we have restructured the manuscript and separated the real-time navigation methodology, the real-time positioning results (Minimum Bounce Criteria), and the post-processing re-positioning improvement (Nearest Bounce Criteria). 
Lastly, when evaluating our methodology, we have removed qualitative language like ``GPS-like'' or ``rivaling GPS'' in favor of quantitative descriptions of ranging error.

We also more prominently establish definitions for navigation, positioning, and re-positioning to better contextualize our results and contributions, based on this discussion in Howe et al (2019) \emph{Observing the Oceans Acoustically} (and many others):

\begin{quote} 
PNT is a combination of three distinct, constituent capabilities:

@@

	By comparing the errors between these two pseudorange estimation methods, we noticed a surprising underlying pattern compared to the ground truth GPS data.
	Further investigation suggests there is noise in the GPS measurement that is not incurred in the travel time measurement that informs the acoustic range estimation.

	The manuscript has been significantly restructured to separate analyses along this narrative.

	\reviewer{}{The hypothesis is stated in the discussion section lines 515 and following: “We hypothesize and validate that the embedded stochastic prediction of a single group velocity is a smoothly varying function of range, source, and receiver depth pairings as well as multipath structure.”  I am not convinced that the few short fixed ranges presented here validate this hypothesis.  Lines 248 and following indicate how the system should track a continuously changing group velocity, but this is not demonstrated with the few fixed ranges presented here.  Also, it would be good to state this hypothesis more clearly earlier in the manuscript.}
	We agree that a few short fixed ranges do not validate this hypothesis. With the overarching motivation and design of LBL navigation, we have recast this hypothesis as an assumption. We would note, again, that the vehicle navigation performance would better determine a continuously changing group velocity, but this data would have no way of being verified.

	\reviewer{}{The introduction and conclusion sections should be tightened up to clarify the state of the art in various arenas of underwater vehicle navigation and how this work fits in.

	The error metrics presented here should be specifically compared with other positioning systems or algorithms over similar spatial scales that do not integrate the ray propagation physics.

	Various ranging and positioning methods are referenced, but the narrative jumps between different types of AUVs and implementations and between scales of a few kilometers and hundreds of kilometers. This should be discussed in terms of range, vehicle, and propagation conditions.

	Line 176 mentions that the SSP and the physics of ray refraction are often overlooked in positioning algorithms; this is, I presume, referring to real-time results or short-range results, but doesn’t specifically say this.}

	We have re-written the introduction and conclusion to clarify that our work adds a necessary model-aided component for ranging in LBL networks for complex acoustic environments, and that it could be extended for long range propagation as well.

	Present error metrics in reference to ...

	The suggestion to discuss in terms of range, vehicle, and propagation conditions is helpful, and we have included that framing explicitly in the manuscript (LINE LINE). However, considering a union of all those operational variables, there are no comparable deployments.
	Previous Arctic deployments are either in the Marginal Ice Zone at late summer, with little to no ice, or recorded the classical Arctic upward refracting profile, where isovelocity scaling could be good enough.
	The few Arctic deployments that have seen a double ducted environment were on gliders, with little to no sea ice, and only one study navigated and re-navigated the gliders acoustically; other studies navigated them with dynamic flight models and GPS and re-navigated them acoustically.

	Lastly, we have removed the line in question, but would point out it was originally written as ``navigation'', not ``positioning''.

@@

	\reviewer{}{Related to the point above, the introduction should make clear what type of vehicles the proposed method of navigation is applicable to and over what ranges. The authors do start to do this when they note that a shallow receiver depth enables the estimation method (line 274), but should more directly specify under what conditions the proposed algorithm is applicable.  In the Discussion section, there is an implication that this method of using a single group velocity could be applied to mesoscale operations but should acknowledge that propagation at these larger scales results in arrivals that are spread over several seconds including multiple modal arrivals with varying group velocities.  This method would presumably result in errors much higher than 11m….how might that scale?}

	\reviewer{}{I understand what the authors are referencing when they use the term “group velocity” but I think this term should be used a bit more carefully.  Group velocities are often associated with modal propagation in which each mode has a distinct group velocity.  The goal in this work is to determine a single group velocity by taking a power weighted average of several different ray arrivals to derive range.  This is addressed somewhat in lines 395 and following, but I think it would be appropriate to expand on this discussion and move it up.  Lines 235 and following state “the variation of horizontal group velocity from any source-receiver pair is the fundamental challenge to implement GPS-like navigation in an LBL navigation paradigm…” this statement could use some clarification.  Also, perhaps change the axis on Figure 5 to read something like ‘implied group velocity.”}

	``It is the variation of group speed from source to source in underwater navigation that is the fundamental challenge to implement a “GNSS-like” navigation system in the ocean with acoustics.''

### 2021-12-22 manuscript/main.tex
hi

\section{Introduction}
\label{sec:1}  
Autonomous underwater vehicles (AUVs) are increasingly capable platforms to explore and samlpe the ocean, particularly for remote and/or dangerous environments.
However, navigation uncertainty is a major challenge in considering AUVs as reliable and standard tools for oceanographic research.
Compared to their land and air-based counterparts, underwater vehicles face a unique set of challenges brought on by the water column itself:  while land and air-based robots utilize the Global Navigation Satellite System (GNSS) to achieve stunning location accuracy and precision throughout the duration of their missions, AUVs cannot access GPS while underwater due to the rapid attenuation of electromagnetic waves \citep{preisig_acoustic_2007}.
Typically, an AUV can stall on the surface to receive a GPS fix, but this foolproof method of re-positioning is inaccessible in an ice-covered environment or other GPS-denied situations.

% Long baseline navigation
Though there many acoustic positioning methods, the long baseline (LBL) is closest in style

% A recent study tracking 22 floats under the ice in the Weddell Gyre near Antarctica showed that they were unable to surface for 8 months and position uncertainty grew to 116 $\pm$ 148 km in the lateral directions \citep{chamberlain_observing_2018}.

Much progress has been made in underwater vehicle navigation, which relies on any combination of dead reckoning, hydrodynamic models, inertial navigation systems, doppler velocity logs, and/or acoustic baseline navigation systems \citep{paull_auv_2014}.
In particular, the long baseline (LBL) navigation system is the most similar in scale and style to GPS, and most appropriate for mitigating drift in underwater vehicle localization without overburdening computation on the vehicle \citep{van_uffelen_global_2021}.

@@

Underwater acoustic navigation (and GPS) is predicated on timing and positioning.
Timing is the ability to maintain accurate and precise time, sometimes synchronized across multiple nodes.
\llabel{1.11} The advent of the WHOI micro-modem \citep{singh_underwater_2006} and synchronized chip scale atomic clocks \citep{gardner_second_2016} enabled ranging via one-way travel time (OWTT).

Positioning is the ability to accurately determine one's location and is generally done via trilateration.
Navigation assumes a degree of autonomy to move from a current position to a desired one.
The standard for real-time navigation efforts assume a locally homogeneous sound speed profile (SSP) that enables a linear, deterministic scaling between travel time and range \citep{eustice_recent_2006,eustice_experimental_2007,webster_preliminary_2009,webster_advances_2012}.
Short range (less than ten kilometers) navigation efforts achieve minimal navigation error with a nominal sound speed value \citep{eustice_experimental_2007,webster_preliminary_2009,kepper_mems_2017}.
For mesoscale (tens to hundreds of kilometers) navigation efforts, a deterministic \citep{graupe_preliminary_2019} or through-the-sensor \citep{webster_towards_2015} value for sound speed is used, but error remains in the hundreds of meters.

Post-processing localization efforts (re-navigation for a self-determinedly moving object or re-positioning for a passively moving one), without computational or time constraints, can indulge more complex acoustic methods and/or ocean representations.
There are many different approaches, from considering a temporally averaged mean depth dependent SSP \citep{van_uffelen_localization_2016}, including spatio-temporal variability in the sound speed profile to minimize error \citep{graupe_preliminary_2019, mikhalevsky_deep_2020}, comparing acoustic records with synthetic acoustic records computed through an ocean model \citep{wu_deep_2019}. 
The ``cold start'' algorithm developed by \citet{mikhalevsky_deep_2020} isolates a specific ray path to find a representative group velocity and re-positioned a floating hydrophone array with an error of 85 m and a standard deviation of 32 m relative to ground truth GPS data, but implementing this algorithm for navigation remains to be seen.
These post-processing efforts have generally been deemed too computationally intensive and/or labor intensive for real-time efforts, where automation of identifying time arrivals is further complicated by multipath.

@@

Underwater navigation research is broadly motivated by GPS-like navigation in GPS-denied conditions.
The Arctic, while remote, is the perfect place to test mature navigation technologies in real GPS-denied conditions.

Literature and historical reviews of underwater navigation approaches in under-ice conditions highlight the emerging technology learning curve for vehicle operations in the Arctic, the hostile conditions brought on by a semi-permanent moving ice cover, and the need for both absolute and relative navigation with little to no tolerance for error \citep{norgren_unmanned_2014,mcfarlane_autonomous_2015,barker_scientific_2020}.

@@

@@

>>>>

\section{Introduction}
\label{sec:1}  
Autonomous underwater vehicles (AUVs) are increasingly capable platforms to explore and sample the ocean, particularly for remote and/or dangerous regions.
However, navigation uncertainty is a major challenge in considering AUVs as standard tools for oceanographic research.
While land and air-based robots utilize information from  Global Positioning System (GPS) updates\textemdash or more broadly, any Global Navigation Satellite System (GNSS)\textemdash to achieve stunning location accuracy and precision throughout the duration of their missions, AUVs cannot access GPS while underwater due to the rapid attenuation of electromagnetic waves.
Therefore, underwater vehicle navigation has relied on any combination of dead reckoning, hydrodynamic models, inertial navigation systems, doppler velocity logs, and acoustic baseline positioning systems \citep{paull_auv_2014}.
Limiting navigation error and drift requires an AUV to periodically stall on the surface and obtain a GPS fix to reset its position error.
This foolproof method of self-positioning is undesirable for stealth, adverse weather conditions, and mission efficiency, and inaccessible in a GPS-denied situation like an ice-covered environment.

% introduce LBL and precise statement of work
Of acoustic baseline positioning systems, long baseline (LBL) is the most GPS-like in style and scale, and most appropriate for mitigating drift in underwater vehicle localization without overburdening computation or payload size on the vehicle \citep{paull_auv_2014,van_uffelen_global_2021}.
The state-of-the-art for LBL assumes an isovelocity, linear scaling between one way travel time (OWTT) and range \citep{eustice_recent_2006,eustice_experimental_2007,webster_preliminary_2009,webster_advances_2012}.
This assumption is valid for short scale operations but introduces error in complex acoustic propagation environments.
Thus, this paper demonstrates embedded acoustic model-aided data processing to convert recorded OWTTs into pseudoranges in real-time.
This approach was necessitated by a low risk tolerance for navigation error, given total under-ice conditions and double ducted acoustic environment, in the Beaufort Sea, in March 2020, during the Ice Exercise 2020 (ICEX20).
We quantify the success of the proposed acoustic ranging method on GPS-linked beacon-to-beacon communication events, not the AUV-LBL positioning itself, as the latter has no ground truth GPS to compare to.

A

@@

In the interest of clarity, we formally reiterate definitions of timing, positioning, and navigation (CITE).
Timing is the ability to maintain accurate and precise time synchronized across multiple nodes.
Positioning is the ability to accurately determine one's location referenced to a standard geodetic system. It can be done via trilateration or triangulation.
Navigation, implying a degree of autonomy, is the ability to distinguish between current and desired position, and apply necessary corrections to course, orientation, and/or speed to attain that desired position.
Thus, navigation is also intrinsically real-time and is predicated on positioning, which is predicated on timing.

Given the computational and physical barriers of assessing underwater position in real-time, we define post-processing counterparts\textemdash re-navigation and re-positioning\textemdash which may be supplemented by additional data, model output, and computationally or labor intensive algorithms.

\subsection{Background on underwater navigation, re-navigation, and re-positioning}

Underwater navigation research is broadly motivated by acquiring GPS-like navigation accuracy (on the order of meters) in GPS-denied conditions.

Navigation efforts are currently limited to small-scale deployments, and take advantage of 
\llabel{1.11} the advent of the WHOI micro-modem \citep{singh_underwater_2006} and synchronized chip scale atomic clocks \citep{gardner_second_2016}.
Short range (less than ten kilometers) navigation efforts achieve minimal error with a nominal sound speed value \citep{eustice_experimental_2007,webster_preliminary_2009,kepper_mems_2017}.
However, these efforts were all conducted with X sound speed conditions...

Of particular note are previous efforts in the Arctic's Beaufort Lens.
At larger scales on the order of hundreds of kilometers, a deterministic \citep{graupe_preliminary_2019} or through-the-sensor \citep{webster_towards_2015} value for sound speed was used, but error remained in the hundreds of meters.

The LBL system provides navigation aiding as position updates to correct for localization drift and error.

It is important to note that post-processing localization efforts (re-navigation for a self-determinedly moving object or re-positioning for a passively moving one), are often applied for large-scale deployments to better account for spatial variability in the ocean state.

Recently, buoyancy gliders have been re-navigated by considering a temporally averaged depth dependent SSP and CTD observations of...

A well known example of long range re-positioning are RAFOS drifters that record the timestamps of the four strongest pings; transmit all data after mission completion; and are manually localized with uncertainties of 1.5 to 4.5 km (CITE ROSSBY, DUDA).

Other re-positioning methods employ four-dimensional ocean models.
\citet{wu_deep_2019} compares acoustic records with synthetic acoustic records computed through HYCOM \cite{chassignet_hycom_2007}.
\citet{mikhalevsky_deep_2020} introduces an impressive ``cold start'' algorithm that isolates a specific ray path to find a representative group speed, re-positioning a floating hydrophone array with an error of 85 m and a standard deviation of 32 m relative to ground truth GPS data.
Implementing these algorithms for navigation remains to be seen.

@@

U
We refer the reader to excellent reviews of vehicle operations in the Arctic by \citet{norgren_unmanned_2014} and \citet{barker_scientific_2020}.
Previous AUV deployments in the Arctic \cite{bellingham_auv_1993}
Literature and historical reviews of underwater navigation approaches in under-ice conditions highlight the emerging technology learning curve for vehicle operations in the Arctic, the hostile conditions brought on by a semi-permanent moving ice cover, and the need for both absolute and relative navigation with little to no tolerance for error \citep{norgren_unmanned_2014,mcfarlane_autonomous_2015,barker_scientific_2020}.

@@

During ICEX20, we deployed AUV \emph{Macrura}, a custom Bluefin-21, in complete under-ice conditions and a complex double ducted environment brought upon by a persistent warm water intrusion in the Pacific Summer Water layer, which we call the Beaufort Lens.
The majority of the navigation updates were below 11 m and were small enough to enable an emergency recovery. 
When considering operational variables such as range, vehicle class, and propagation conditions, there are no comparable deployments.

@@

Underwater navigation research is broadly motivated by acquiring GPS-like navigation accuracy in GPS-denied conditions.
The Arctic, while remote, is the perfect place to test mature navigation technologies in real GPS-denied conditions.
